---
title: "Semesteroppgave SOK-1005-v22"
author: "Kandidatnr 9"
date: "7 Juni 2022"
output: 
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    number_sections: true
  
editor_options: 
  markdown: 
    wrap: 72
---

Loading packages:

```{r downloading packages}
library(tidyverse)
library(lubridate)
library(knitr)

```

## **Oppgave 1: Slå sammen alle datasettene**

<br/> **Downloading data**

```{r downloads}
file_location <- paste0(getwd(), '/')

AppWichStoreAttributes <-
  "https://data.mendeley.com/public-files/datasets/6htjnfs78b/files/26afd5e7-90b1-4301-ac5e-5905b38c4ec2/file_downloaded"
dest <- paste0(file_location, "AppWichStoreAttributes.csv")
download.file(AppWichStoreAttributes, dest)

county_crime <-
  "https://data.mendeley.com/public-files/datasets/6htjnfs78b/files/3691994e-2874-4ee6-b215-12e130c96175/file_downloaded"
dest <- paste0(file_location, "county_crime.csv")
download.file(county_crime, dest)

county_demographic <-
  "https://data.mendeley.com/public-files/datasets/6htjnfs78b/files/527e7486-7233-460a-99e0-3529b7cd7d49/file_downloaded"
dest <- paste0(file_location, "county_demographic.csv")
download.file(county_demographic, dest)

county_employment <-
  "https://data.mendeley.com/public-files/datasets/6htjnfs78b/files/846ac757-721e-4fd9-a414-5871f60de093/file_downloaded"
dest <- paste0(file_location, "county_employment.csv")
download.file(county_employment, dest)

WEEKLY_SALES_10STORES <-
  "https://data.mendeley.com/public-files/datasets/6htjnfs78b/files/b963fdd1-0df9-4257-bd62-65256ec9d57c/file_downloaded"
dest <- paste0(file_location, "WEEKLY_SALES_10STORES.csv")
download.file(WEEKLY_SALES_10STORES, dest)

WEEKLY_WEATHER <-
  "https://data.mendeley.com/public-files/datasets/6htjnfs78b/files/b8708b3c-c5e8-456e-83f4-9f23f472a840/file_downloaded"
dest <- paste0(file_location, "WEEKLY_WEATHER.csv")
download.file(WEEKLY_WEATHER, dest)

# Removes all noise since we have finished downloading
rm(list=ls()) 
```

- **I change the names of the datatables:**
  - AppWichStoreAttributes = df1_attr
  - county_crime = df2_crime
  - county_demographic = df3_demo
  - county_employment = df4_em
  - weekly_weather = df5_weather
  - weekly_sales_10stores = df6_sales

<br/> 
**I go through the dfs and make necessary changes in names and
formats:**

```{r renaming and going through the dfs}

# some changes
# change the names of Store_County and Store_Weather_Station
df1_attr <- AppWichStoreAttributes <- read_csv("AppWichStoreAttributes.csv")
df1_attr_2 <-df1_attr %>%
  rename(County_Name = Store_County, Weather_Station = Store_Weather_Station)
head(df1_attr_2)


# no changes:
df2_crime <- county_crime <- read_csv("county_crime.csv")
head(df2_crime)


# no changes:
df3_demo <- county_demographic <- read_csv("county_demographic.csv")
head(df3_demo)


# no changes:
df4_em <- county_employment <- read_csv("county_employment.csv")
head(df4_em)


# some changes:
# Formatting the weather date:
df5_weather <- WEEKLY_WEATHER <- read_csv("WEEKLY_WEATHER.csv")

df5_weather$Weather_Date <- 
  as.Date(df5_weather$Weather_Date, format = "%d/%m/%Y") 
head(df5_weather)


# Some changes:
# rename the store number column
# some dates are missing in the Date column, I use the other date columns.
# removing unnecessary columns
# changing the letters in Description (names of items)
df6_sales <- WEEKLY_SALES_10STORES <- read_csv("WEEKLY_SALES_10STORES.csv")
df6_sales_2 <- df6_sales %>% 
  rename(Store_Num = Store_num) %>% 
  mutate(Date = as.Date(with(df6_sales, paste(Year, Month, Day,sep="-")), "%Y-%m-%d"))

df6_sales_3 <- 
  subset (df6_sales_2, select = -c(Year, Month, Day)) 
df6_sales_3$Description = str_to_title(df6_sales_3$Description)
df6_sales_3

```

**I make one df out of alle df by using the function left_join():**

```{r merging dfs}
df1_df2 <- left_join(df1_attr_2, df5_weather, by = "Weather_Station")
df1_df3 <- left_join(df1_df2, df2_crime, by = "County_Name")
df1_df4 <- left_join(df1_df3, df3_demo, by = "County_Name")
df1_df5 <- left_join(df1_df4, df4_em, by = "County_Name")
All_df <- left_join(df1_df5, df6_sales_3, by = c("Store_Num", "Weather_Date" = "Date"))

# renaming weather_Date and filtering out Cost <= 0
All_df <- All_df %>% 
  rename("Date" = "Weather_Date") %>% 
  filter(Cost > 0) 
All_df
```

```{=html}
<pre>

</pre>
```
## **Oppgave 2: Lag en ukesrapport for et utsalg**

```{r links_to_name_and_period}
# Store Name
Store_No_Nm <- All_df %>% 
  filter(Store_Num == 14)  

# Week number
Week <- All_df %>% 
  filter(Weather_Week == 20) 
```

This is a weekly report for store no `r Store_No_Nm[1,2]`
`r Store_No_Nm[1,1]` for the period week no `r Week[1,22]` which gives a
overview and assessment of the results.

A weekly sales report will be a status report of how the operations in a
shop are going. Key issues will be how sales of various goods have
been and on which goods the Profit has been greatest.
The dataset includes many types of items. It has therefore been chosen
to divide the product range into price groups and look at which price
group has the largest Sale and Profit. Here it is also looked at which
item is sold the most within the price group with the largest Sales. It
has also been chosen to look at goods with the lowest sales, and this
can be helpful in assessing whether to maintain the large breadth of the
product range.
It must be added that the Profit is Gross Profit so that there will not
be a supplementary analysis.

```{r selecting variables for week report}
# selecting week and Store
sales_14_20 <-All_df %>% 
  filter(Weather_Week == 20, Store_Num == 14) %>% 
  select(Store_Name, Store_Num, Store_City, County_Name, Date, INV_NUMBER,Description, Price, Sold, 
         Sales, Tot_Sls, Unit_Cost, Cost,Cost_Percent, Margin, Profit) %>% 
  group_by(INV_NUMBER, Description, Price, Sold, Cost, Profit, Margin) %>% 
  summarise(Sales) %>% 
  ungroup()
```

```{r item most sales}
# Item that have most Sale
sales_most_item <-
  subset(sales_14_20, Sales = max(sales_14_20$Sales)) %>% 
  arrange(desc(Sales))
sales_most_item

```

```{r item most profitable}
Profit_most_item <-
subset(sales_14_20, Profit = max(sales_14_20$Profit)) %>% 
  arrange(desc(Profit))
Profit_most_item
```

```{r comparing two weeks}
# selecting the week before to compare development in Sales and Profit:
sales_14_19 <-All_df %>% 
  filter(Weather_Week == 19, Store_Num == 14) %>% 
  select(Store_Name, Store_Num, Store_City, County_Name, Date, INV_NUMBER,Description, Price, Sold, 
         Sales, Tot_Sls, Unit_Cost, Cost,Cost_Percent, Margin, Profit) %>%
  group_by(INV_NUMBER, Description, Price, Sold, Cost, Profit, Margin) %>% 
  summarise(Sales) %>% 
  ungroup()

# sum of variables:
sum(sales_14_19$Sales)
sum(sales_14_20$Sales)
sales_change = round(100 * ((sum(sales_14_20$Sales)/sum(sales_14_19$Sales)-1)),1)
sales_change

sum(sales_14_19$Profit)
sum(sales_14_20$Profit)
profit_change = round(100 * ((sum(sales_14_20$Profit)/sum(sales_14_19$Profit)-1)),1)
profit_change
```

**Comment**\
Total revenues of Sales for store `r Store_No_Nm[1,2]`
`r Store_No_Nm[1,1]` i week `r Week[1,22]` were \$
`r sum(sales_14_20$Sales)`. The total sales changed `r sales_change`
percent compared to the week before.\
Total profit in the same period were \$ `r sum(sales_14_20$Profit)`. The
change in total profit from the week before were `r profit_change`
percent.\
Sales revenues were greatest for `r sales_most_item[1,2]`. Total sales
for this were \$ `r sales_most_item[1,8]`. The most profitable item were
`r Profit_most_item[1,2]`. For this, the total profit were \$
`r Profit_most_item[1,6]`.

<br/>

### **Sorting the goods after price groups**

Due to the large breadth of the product range, it has been chosen to
take the product prices as a starting point for the analysis and group
them. They are sorted in following groups. Price below \$1(here are also
refund etc that is negative figures), between \$1 and \$2, and so on.
The last price group is price over \$8.

```{r sorting price in groups}
sales_price_gr <- sales_14_20 %>% 
  group_by(price_group = ifelse(Price <= 1.0, "price_below_$1", 
                         ifelse(Price > 1 & Price <= 2, "price_$1", 
                         ifelse(Price > 2 & Price <= 3, "price_$2", 
                         ifelse(Price > 3 & Price <= 4, "price_$3", 
                         ifelse(Price > 4 & Price <= 5, "price_$4", 
                         ifelse(Price > 5 & Price <= 6, "price_$5", 
                         ifelse(Price > 6 & Price <= 7, "price_$6", 
                         ifelse(Price > 7 & Price <= 8.0, "price_$7", "price_over_$8"))))))))) %>% 
  summarise(Sold, Price, Sales, Profit)

# we remove all negative values from the dataset
sales_price_gr <-
  sales_price_gr %>%
  select(price_group, Sold, Price, Sales, Profit) %>%
  filter(price_group >= 0) %>%
  group_by(price_group) %>%
  summarise(Tot_sales = sum(Sales), Tot_sold = sum(Sold), Tot_profit = sum(Profit)) 

# Total sales and profit:
sum(sales_price_gr$Tot_sales)
max(sales_price_gr$Tot_sales)  

sum(sales_price_gr$Tot_profit)
max(sales_price_gr$Tot_profit)


sales_most <- 
  subset(sales_price_gr, Tot_sales == max(sales_price_gr$Tot_sales))


profit_most <- 
  subset(sales_price_gr, Tot_profit == max(sales_price_gr$Tot_profit))
```

### **Share of sales and profit per pricegroup**

```{r shares of sales and profit}
# share of sales and profit per pricegroup:
sales_share <-
  sales_price_gr %>% 
  mutate(Share_sales = round(100 *(sales_price_gr$Tot_sales/sum(sales_price_gr$Tot_sales)), 1),
         Share_profit = round(100 *(sales_price_gr$Tot_profit/sum(sales_price_gr$Tot_profit)),1)) %>% 
  select(price_group, Tot_sold, Tot_sales, Share_sales, Tot_profit, Share_profit)

Shares_most <- 
  subset(sales_share, Share_sales == max(sales_share$Share_sales)) 
```

```{r table 1 Amount and shares for Sale and Profit}
Goods_shares <- sales_share[1:9, ]
  kable(Goods_shares, caption = "Table 1: Amount and shares for Sale and Profit")
```

**Comment**\
We see from table 1 the highest revenue comes from sale in the
pricegroup `r Shares_most[1,1]`. The sales here are \$
`r Shares_most[1,3]` and profit are \$ `r Shares_most[1,5]`. This
pricegoup has a share of `r Shares_most[1,4]` percent of total Sale and
`r Shares_most[1,6]` percent of total Profit.

<br/>

#### **Figur 1 plot of sales per pricegroup**

```{r figure 1 sales salsgroup}
figure_1 <-
sales_price_gr %>% 
ggplot(aes(x=price_group, y = Tot_sales))+
  geom_bar(stat= "identity", fill = "steelblue") +
  geom_text(aes(label=Tot_sales), vjust= -0.3, size=3.5)+
  labs(title = "Figure 1: Sales per pricegroup", x = "Pricegroups in $", y = "Total Sales per pricegoup",
        caption = "Figure 1 shows how Sales are distributed across pricegroups. For example' price_$5 mean goods that cost from $4 to $5.") +
  theme_classic()
figure_1

```

**Comment**\
We see from Figure 1 that the highest sales are of goods in the price
group `r sales_most[1,1]`. Total Sales are \$`r sales_most[1,2]`. The
first price group has negative sales figures. The reason for this is
that it also contains refunds etc.\
<br/>

#### **Figure 1a plot of shares of sales per pricegroup**

```{r figure 1a share of sales}
figure_1a <-
  sales_share %>% 
  ggplot(aes(x=price_group, y = Share_sales))+
  geom_bar(stat= "identity", fill = "steelblue") +
  geom_text(aes(label=Share_sales), vjust= -0.3, size=3.5)+
  labs(title = "Figure 1a: Shares of Sales per pricegroup", x = "Pricegroups in $", y = "Share of Sales per pricegoup in percent",
       caption = "Figure 1 shows how shares of Sales are distributed across pricegroups.") +
  theme_classic()
figure_1a
```

**Comment**\
We see from Figure 1a that the highest share of sales are of goods in
the price group `r Shares_most[1,1]`. This price group has
`r Shares_most[1,4]` per cent of all sales.\
<br/>

#### **Figure 2 plot of profit per pricegroup:**

```{r figure 2 profit}
figure_2 <-
  sales_price_gr %>% 
  ggplot(aes(x=price_group, y = Tot_profit))+
  geom_bar(stat= "identity", fill = "steelblue") +
  geom_text(aes(label=Tot_profit), vjust= -0.3, size=3.5)+
  labs(title = "Figure 2: Profit per pricegroup", x = "Pricegroups in $", y = "Total Profit per pricegoup",
       caption = "Figure 2 shows how Profit are distributed across pricegroups. For example price_$5 mean goods that cost from $4 to $5.") +
  theme_classic()
figure_2
```

**Comment**\
We see from Figure 2 that the largest profit is achieved from sales in
the price group `r profit_most [1,1]`. The total profit here is \$
`r profit_most [1,4]`. The first price group is negative. The reason for
this is that it also contains refunds etc. who will negatively affect
profits.\
<br/>

### **The best-selling items**

```{r best selling item}
# Highest number of items sold
Goods_most<- sales_14_20 %>% 
  filter(Sold > 10) %>% 
  arrange(desc(Sold))

# Higest revenue from Sales
Goods_most_sales<- sales_14_20 %>% 
  filter(Sold > 10) %>% 
  arrange(desc(Sales))

# Higest Profit
Goods_most_p <- sales_14_20 %>% 
  filter(Sold > 10) %>% 
  arrange(desc(Profit))

sum(Goods_most$Sold)
sum(Goods_most$Sales)
sum(Goods_most$Profit)

```

**We make a table of most sold items**

```{r most sold table 2}
Goods_most <- Goods_most[1:10, ]
  kable(Goods_most, caption = "Table 2: Most sold items")
```

**Comment**  

Table 2 shows the 10 items that sell the most in Store
`r Store_No_Nm[1,2]` `r Store_No_Nm[1,1]` in this week. Sales for these
10 are total $ `r sum(Goods_most$Sales)` and account for
`r round(100 *(sum(Goods_most$Sales)/sum(sales_14_20$Sales)),1)` percent
of total sales.

It appears from Table 2 that the mostselling item were
`r Goods_most [1,2]`. `r Goods_most [1,4]` units of `r Goods_most [1,2]`were sold. That account for
$`r Goods_most[1,8]` in sale. The sale of `r Goods_most[1,2]` accounts for $`r round(100 *(Goods_most[1,8]/sum(sales_14_20$Sales)),1)` percent of total Sales.

In second place we find `r Goods_most[2,2]` with `r Goods_most[2,4]`
units and \$ `r Goods_most[2,8]` of sale. As third ranked, we find
`r Goods_most[3,2]` with `r Goods_most[3,4]` units and \$
`r Goods_most[3,8]` of sale.

`r Goods_most_p[1,2]` is the most profitable sold item. The profit is \$
`r Goods_most[1,6]`. This amounts to
`r round(100 *(Goods_most_p[1,6]/sum(sales_14_20$Profit)),1)` percent of
total profit.
<br/>

### **The lowest sales**

```{r low sales}
# lowest number of items sold
Goods_low<- sales_14_20 %>% 
  filter(Sold > 10, Price > 0) %>% 
  arrange(Sold)

# lowest revenue from Sales
Goods_low_sales<- sales_14_20 %>% 
  filter(Sold > 10, Price > 0) %>% 
  arrange(Sales)

# lowest Profit
Goods_lowest_p <- sales_14_20 %>% 
  filter(Sold > 10, Price > 0) %>% 
  arrange(Profit)
Goods_lowest_p
```

**We make a table of the lowest number of sold items**

```{r low sales table 3}
Goods_low <- Goods_low[1:10, ]
  kable(Goods_low, caption = "Table 3: lowest number sold items")
```

**Comment**
From table 3 we see that there are many items that not are selling
good.
<br/>

### **Summary**

The week sales report for Store `r Store_No_Nm[1,2]`
`r Store_No_Nm[1,1]` in week `r Week[1,22]`show that total revenues for
were \$ `r sum(sales_14_20$Sales)`, and total Profit were \$
`r sum(sales_14_20$Profit)`.\
The most-selling item were `r sales_most_item[1,2]`. Total sales for
this were \$ `r sales_most_item[1,8]`. The total sales changed
`r sales_change` percent compared to the week before.\
The most profitable item were `r Profit_most_item[1,2]`. For this, the
total profit were \$ `r Profit_most_item[1,6]`. The change in total
profit from the week before were  `r profit_change` percent.

```{=html}
<pre>

</pre>
```
## **Oppgave 3: Lag en månedlig salgrapport på aggregert nivå.**

<br/>

### **Introduction**

In a monthly report, the group management will be able to see the total
results of the business and make strategic assessments based on this.
This means that Sales and Profits in total and per Store are central. On
the basis of this, it will be possible to assess whether special
measures are necessary for some Stores. Comparison of data on the same
item across Stores will also be relevant. This can form the basis for
looking at whether the product range should be the same in all Stores.

It has been chosen here to divide the product range into price groups
and look at which price group has the largest Sales and Profits. Here it
is also looked at which item is sold the most within the price group
with the largest Sales.

It has also been chosen to look at items with the lowest sales, which
can be helpful in assessing whether to maintain the large breadth of the
product range.

The dataset is not very suitable for creating real monthly data because
it is based on weeks. It will be the case that a week can contain data
for several months. It is therefore recommended that the group switch to
a day-based accounting system so that monthly data can be created in a
simpler way.
<br/>

### **Cleaning the data set for unnecessary variable and grouping some variables**

```{r cleaning of df}

All_df_rev <- All_df %>% 
  
# We merge small population groups into one variable:
  mutate("All_Other_Groups"= rowSums(All_df[ , c("County_Non-Hispanic_Black","County_Non-Hispanic_Asian",
                                                 "County_Non-Hispanic_Pacific_Islander", "County_Non-Hispanic_Two_or_more",
                                                 "County_Hispanic_Black", "County_Hispanic_Asian",
                                                 "County_Hispanic_Pacific_Islander", "County_Hispanic_Two_or_more")])) %>%  
  
# and select the relevant variables:    
  select(Store_Name, Store_Num, Store_City, County_Name, Weather_Station, Date, Weather_Week,    
         Description, Price, Sold,Sales, Tot_Sls, Cost, Margin, Profit, Annual_Rent_Estimate,Store_Location, Store_Drive_Through, 
         Store_Competition_Fastfood, Weather_Bad_Weather_days, County_Total_Crime_Rate, County_Violent_Rate, 
         County_Property_Rate, County_Society_Rate, County_Unemployment_Rate, `County_Non-Hispanic_White`,
         `County_Non-Hispanic_Native_American`, County_Hispanic_White,  County_Hispanic_Native_American, All_Other_Groups, 
         County_Total_Census_Pop, ) %>% 
  
  
# making percentages of population groups
  mutate(All_Other_Groups_pct = round(100 * (All_Other_Groups/County_Total_Census_Pop),1),
         `County_Non-Hispanic_White_pct` = round(100 * (`County_Non-Hispanic_White`/County_Total_Census_Pop),1),
         `County_Non-Hispanic_Native_American_pct` = round(100 * (`County_Non-Hispanic_Native_American`/`County_Total_Census_Pop`),1),
         `County_Hispanic_White_pct` = round(100 * (`County_Hispanic_White`/County_Total_Census_Pop),1),
         `County_Hispanic_Native_American_pct` = round(100 * (`County_Hispanic_Native_American`/County_Total_Census_Pop),1)) %>% 
  
# removing the variables we don't use
  select(- c(`County_Non-Hispanic_White`,
             `County_Non-Hispanic_Native_American`, County_Hispanic_White,  County_Hispanic_Native_American, All_Other_Groups, 
             County_Total_Census_Pop)) %>% 
  rename(Week_No = Weather_Week) 


All_df_rev$County_Violent_Rate <-
  round(All_df_rev$County_Violent_Rate,1)
```

**We choose 4 weeks from 40 to 43:**

```{r choosing weeks}
# choosing weeks
sales_mnd <-All_df_rev %>% 
  filter(Week_No >= 40, Week_No <= 43) %>% 
  select(Store_Name, Store_Num, Store_City, County_Name, Date,Description, Price, Sold, 
         Sales, Profit) %>% 
  group_by(Description, Price, Sold, Sales,Profit) %>% 
  summarise(Sold, Sales) %>% 
  ungroup()
sales_mnd

# sum of variables:
sum(sales_mnd$Sales)
sum(sales_mnd$Profit)
```

**Comment**\
The total Sales revenue from all the Stores in this month were
$`r (sum(sales_mnd$Sales))/1000`. Total Profit amounts to $`r
(sum(sales_mnd\$Profit))/1000\`. All numbers are in thousands of \$.\
<br/>

### **Sorting the goods after price groups**

Due to the large breadth of the product range, it has been chosen to
take the product prices as a starting point for the analysis and group
them. They are sorted in following groups. Price below \$1(here are also
refund etc that is negative figures), between \$1 and \$2, and so on.
The last price group is price over \$8.

```{r sorting pricegroups for use on aggregate level}
sales_price_gr_mnd <- sales_mnd %>% 
  group_by(price_group = ifelse(Price <= 1.0, "price_$1", 
                         ifelse(Price > 1 & Price <= 2, "price_$2", 
                         ifelse(Price > 2 & Price <= 3, "price_$3", 
                         ifelse(Price > 3 & Price <= 4, "price_$4", 
                         ifelse(Price > 4 & Price <= 5, "price_$5", 
                         ifelse(Price > 5 & Price <= 6, "price_$6", 
                         ifelse(Price > 6 & Price <= 7, "price_$7", 
                         ifelse(Price > 7 & Price <= 8.0, "price_$8", "price_over_$8"))))))))) %>% 
  summarise(Description, Sold, Price, Sales, Profit)
sales_price_gr_mnd
```

grouping after pricegroups and items:

```{r grouping by pricegroups and items}
# sum pricegroups
price_gr_Sales_mnd <-
  sales_price_gr_mnd %>%
  select(Description, price_group, Sold, Price, Sales, Profit) %>%
  filter(price_group >= 0) %>%
  group_by(price_group) %>%
  summarise(Tot_sales = sum(Sales), Tot_sold = sum(Sold), Tot_profit = sum(Profit)) 
price_gr_Sales_mnd

# sum items
sales_Description_mnd <-
  sales_price_gr_mnd %>%
  select(Description, price_group, Sold, Price, Sales, Profit) %>%
  filter(price_group >= 0) %>%
  group_by(Description, price_group) %>%
  summarise(Tot_sales = sum(Sales), Tot_sold = sum(Sold), Tot_profit = sum(Profit)) 
sales_Description_mnd
```

```{r sum of variables}
# Total sales and profit:
round(sum(price_gr_Sales_mnd$Tot_sales),0)
round(max(price_gr_Sales_mnd$Tot_sales),0) 

round(sum(price_gr_Sales_mnd$Tot_profit),0)
round(max(price_gr_Sales_mnd$Tot_profit),0) 

```

### **Amount and Share of Sales and Profit per price group**

```{r amount and shares of Sales and Profit}
# amount and shares of Sales and Profit per price group
amount_share_pricegroup_mnd <-
  price_gr_Sales_mnd %>% 
  mutate(Share_sales_mnd = round(100 *(Tot_sales/sum(price_gr_Sales_mnd$Tot_sales)), 1),
         Share_profit_mnd = round(100 *(Tot_profit/sum(price_gr_Sales_mnd$Tot_profit)),1)) %>% 
  select(price_group, Tot_sold, Tot_sales, Share_sales_mnd, Tot_profit, Share_profit_mnd) %>% 
  arrange(desc(Tot_sales))
amount_share_pricegroup_mnd

# Amount of Sales
sales_most_pricegroup_mnd <- 
  subset(amount_share_pricegroup_mnd, Tot_sales == max(amount_share_pricegroup_mnd$Tot_sales))
sales_most_pricegroup_mnd

# Share of total Sales
Shares_most_pricegroup_mnd <- 
  subset(amount_share_pricegroup_mnd, Share_sales_mnd == max(amount_share_pricegroup_mnd$Share_sales_mnd)) 
Shares_most_pricegroup_mnd

# Amount of Profit
profit_most_pricegroup_mnd <- 
  subset(amount_share_pricegroup_mnd, Tot_profit == max(amount_share_pricegroup_mnd$Tot_profit))
profit_most_pricegroup_mnd

# Share of total Profit
profit_share_most_pricegroup_mnd <-
  subset(amount_share_pricegroup_mnd, Share_profit_mnd == max(amount_share_pricegroup_mnd$Share_profit_mnd))  
profit_share_most_pricegroup_mnd

# We make a table of amount and shares
amount_shares_pricegroup_mnd <- amount_share_pricegroup_mnd[1:9, ]
kable(amount_shares_pricegroup_mnd, caption = "Table 4: Amount and shares for Sale and Profit per price group")
```

**Comment**\
We see from Table 4 that the highest sales revenue comes from the price
group `r amount_shares_pricegroup_mnd [1,1]`. The revenues were \$
`r (sales_most_pricegroup_mnd [1,3])/1000`. The profit from sales in
this price group was `r (profit_most_pricegroup_mnd [1,5])/1000`. All
numbers are in thousands of \$.

<br/>

```{r Item with highest Sale in price group}
# bestselling items in price group with highest Sale
bestselling_pricegroup_item <- 
  sales_Description_mnd %>% 
  filter(price_group%in%sales_most_pricegroup_mnd) %>% 
  arrange(desc(Tot_sold))


# We make a table of the 10 bestselling items in pricegroup
bestselling_pricegroup_item <- bestselling_pricegroup_item[1:10, ]
kable(bestselling_pricegroup_item, caption = "Table 5: Most sold items in price group ")
```

**Comment**

We see from table 5 that the bestselling item in the price group that
has the highest Sale is `r bestselling_pricegroup_item[1,1]`. In second
place we find `r bestselling_pricegroup_item[2,1]`.

<br/>

### **Amount and Share of Sales and Profit per item**

```{r Amount and Share of Sales and Profit per item}
# Finding the overall bestselling item
amount_share_item_mnd <-
  sales_Description_mnd %>% 
  mutate(Share_sales_mnd = round(100 *(Tot_sales/sum(sales_Description_mnd$Tot_sales)), 1),
         Share_profit_mnd = round(100 *(Tot_profit/sum(sales_Description_mnd$Tot_profit)),1)) %>% 
  select(Description, price_group, Tot_sold, Tot_sales, Share_sales_mnd, Tot_profit, Share_profit_mnd) %>% 
  arrange(desc(Tot_sales))
amount_share_item_mnd
```

```{r Amount and Share of Sales and Profit per item 2}
# Amount of Sales
sales_most_item_mnd <- 
  subset(amount_share_item_mnd, Tot_sales == max(amount_share_item_mnd$Tot_sales))
sales_most_item_mnd

# Share of total Sales
Shares_most_item_mnd <- 
  subset(amount_share_item_mnd, Share_sales_mnd == max(amount_share_item_mnd$Share_sales_mnd)) 
Shares_most_item_mnd 

# Amount of Profit
profit_most_item_mnd <- 
  subset(amount_share_item_mnd, Tot_profit == max(amount_share_item_mnd$Tot_profit))
profit_most_item_mnd

# Share of total Profit
profit_share_most_item_mnd <-
  subset(amount_share_item_mnd, Share_profit_mnd == max(amount_share_item_mnd$Share_profit_mnd))  
profit_share_most_item_mnd

#We make a table of amount and shares
amount_shares_item_mnd <- amount_share_item_mnd[1:10, ]
kable(amount_shares_item_mnd, caption = "Table 6: Amount and shares for Sale and Profit per item")
```

**Comment**

We see from Table 6 that the highest sales revenue comes from the item
`r amount_shares_item_mnd [1,1]`. The Sales revenue were \$
`r (Shares_most_item_mnd [1,4])/1000`. The profit from sales of this
item was `r (profit_share_most_item_mnd  [1,6])/1000`. All numbers are
in thousands of $.
<br/>

#### **Plot of Sales of mostselling items**

```{r plot figure 3 sales revenue per mostselling items}
# Sales pr item
figure_3 <-
  amount_shares_item_mnd %>% 
  ggplot(aes(x= Description, y = Tot_sales))+
  geom_bar(stat= "identity", fill = "steelblue") +
  geom_text(aes(label=Tot_sales/1000), vjust= -0.3, size=3.5)+
  labs(title = "Figure 3: Sales revenue of mostselling items in one month", 
       x = "Item Description", y = "Sales revenue in $1000",
       caption = "Figure 3 shows total revenues from the 10 mostselling items.") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
figure_3
```

**Comment**
We see from Figure 3 that the largest Sales revenue is achieved by
`r sales_most_item_mnd [1,1]`. The sales revenue from this item amounts
to $`r (sales_most_item_mnd[1,4])/1000`. Numbers are in thousands of
$.
<br/>

#### **Plot of Sales of bestselling items as shares of total Sales**

```{r plot figure 3a bestselling items shares}
# Sales as shares of total Sales
figure_3a <-
  amount_shares_item_mnd %>% 
  ggplot(aes(x= Description, y = Share_sales_mnd))+
  geom_bar(stat= "identity", fill = "steelblue") +
  geom_text(aes(label=Share_sales_mnd), vjust= -0.3, size=3.5)+
  labs(title = "Figure 3a: Sales of bestselling items as shares of total Sales in one month", 
       x = "Item Description", y = "Sales shares per item",
       caption = "Figure 3 shows the 10 bestselling item as shares of total Sale .") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
figure_3a
```

**Comment**\
We see from Figure 3a that the largest share of total Sales is achieved
by `r Shares_most_item_mnd [1,1]`. This item has
`r Shares_most_item_mnd [1,5]` percent of total Sales.\
<br/>

#### **Plot of profit on bestselling items**

```{r plot figure 4 profit per item}
# Profit as share of total Profit
figure_4 <-
  amount_shares_item_mnd %>% 
  ggplot(aes(x= Description, y = Share_profit_mnd))+
  geom_bar(stat= "identity", fill = "steelblue") +
  geom_text(aes(label= Share_profit_mnd), vjust= -0.3, size=3.5)+
  labs(title = "Figure 4: Profits of mostselling items as shares of total Profit in one month", 
       x = "Item Description", y = "Profit shares per item",
       caption = "Figure 4 shows the 10 mostselling item as shares of total Profit .") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
figure_4
```

**Comment**\
We see from Figure 4 that the largest profit is achieved from sales in
the price group `r profit_most_item_mnd[1,1]`. This item has
`r profit_most_item_mnd[1,7]` percent of total Profit.\
<br/>

### **Aggregating per Store**

```{r Aggregating per Store}
# making aggregate sum per Store
Store_mnd <- All_df_rev %>% 
  filter(Week_No >= 40, Week_No <= 43) %>% 
   select(Store_Name, Store_Num,Description, Sold, Sales, Profit, Store_Location, Store_Drive_Through, Store_Competition_Fastfood,
          County_Total_Crime_Rate, County_Unemployment_Rate, `County_Non-Hispanic_White_pct`,County_Hispanic_Native_American_pct, 
         County_Hispanic_White_pct,`County_Non-Hispanic_Native_American_pct`, All_Other_Groups_pct) %>% 
  group_by(Store_Num,Store_Name, Store_Location, Store_Drive_Through, Store_Competition_Fastfood,
           County_Total_Crime_Rate, County_Unemployment_Rate, `County_Non-Hispanic_White_pct`,County_Hispanic_Native_American_pct, 
           County_Hispanic_White_pct,`County_Non-Hispanic_Native_American_pct`, All_Other_Groups_pct) %>% 

  summarise(Store_Name_Sold = sum(Sold), Store_Name_Sale = sum(Sales)/1000, Store_Name_Profit = sum(Profit)/1000)

  # making shares
  Store_mnd <- Store_mnd %>% 
  mutate(sold_share = round(100 *(Store_Name_Sold/sum(Store_mnd$Store_Name_Sold)),1),
         sales_share = round(100 *(Store_Name_Sale/sum(Store_mnd$Store_Name_Sale)),1),
         profit_share = round(100 *(Store_Name_Profit/sum(Store_mnd$Store_Name_Profit)),1)) %>% 
  
  # creates the desired order of the variables
  select(Store_Num,Store_Name,Store_Name_Sold, sold_share, Store_Name_Sale,sales_share,
         Store_Name_Profit, profit_share,Store_Location, Store_Drive_Through, Store_Competition_Fastfood,
         County_Total_Crime_Rate, County_Unemployment_Rate, `County_Non-Hispanic_White_pct`,County_Hispanic_Native_American_pct, 
         County_Hispanic_White_pct,`County_Non-Hispanic_Native_American_pct`, All_Other_Groups_pct) %>% 
  rename(other_demo_groups_pct = All_Other_Groups_pct,
         sold = Store_Name_Sold,
         sales = Store_Name_Sale,
         profit = Store_Name_Profit) %>%  
  ungroup()
Store_mnd
```

```{r sum og variables 2}
sum(Store_mnd$sold)
sum(Store_mnd$sales)
sum(Store_mnd$profit)
```

```{r Table 7 Amount and shares for Sold, Sales and Profit per Store }
#We make a table of amount and shares for each Store
amount_shares_store_mnd <- Store_mnd[,1:8 ]
kable(amount_shares_store_mnd, caption = "Table 7: Amount and shares for Sold, Sales and Profit per Store. In $1000.")
```

```{r arranging economic variables}
# max of the economic variables
Store_most_sold <- Store_mnd %>% 
  arrange(desc(sold))

Store_most_sales <- Store_mnd %>% 
  arrange(desc(sales))

Store_most_profit <- Store_mnd %>% 
  arrange(desc(profit))
```

**Comment**\
We see from Table 7 that the store with the largest sales was
`r Store_most_sales[1,2]`. The amount was \$`r Store_most_sales[1,5]` in
thosand \$ and was `r Store_most_sales[1,6]` percent of the total sales
for all the stores.\
The store with the greatest Profit was `r Store_most_profit[1,2]`. It is
calculated to \$`r Store_most_profit[1,7]` in thousand \$. It was
`r Store_most_profit[1,2]` that had the highest share of the total
Profit. `r Store_most_profit[1,2]` share was `r Store_most_profit[1,8]`
percent. It must be emphasized that it is the gross profit that is
calculated here.

### **Plot of Sales and shares per Store**

```{r Figure 5 sales shares per Store}
# sales shares per Store}
figure_5 <-
  amount_shares_store_mnd %>% 
  ggplot(aes(x= Store_Name, y = sales_share))+
  geom_bar(stat= "identity", fill = "steelblue") +
  geom_text(aes(label=sales_share), vjust= -0.3, size=3.5)+
  labs(title = "Figure 5: Sales shares per Store in one month ", 
       x = "Store name", y = "Sales shares in percent",
       caption = "Figure 5 shows sales share of  total Sale per Store .") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
figure_5
```

**Comment**

Figure 5 shows share of total Sales for all the stores. We see that
there are large differences between the stores in the share of total
Sales.

<br/>

### **Correlation between the variables**

```{r Correlation between the non-economic variables}
# Looking at the non-economic variables
non_ec_var <- Store_mnd %>% 
  select(Store_Num,Store_Name, sold,sales,profit, Store_Location, Store_Drive_Through) %>% 
  arrange(desc(sales))
non_ec_var
```

**Comment**\
We see that only the store with the highest sales is located as Free
Standing. This store also has Store_Drive_Through as one of two in the
dataset. These factors can be important framework conditions for
localization because they clearly affect sales.

<br/> **Correlation**\
The relationship between total sales and a number of variables in the
data set has also been calculated. It must be emphasized that this will
only apply to the values ​​in the selected month.

```{r correlations between the variables}

# Preparing the dataset for regression
res_store_mnd <- Store_mnd %>% 
  select(sales,County_Total_Crime_Rate, 
         County_Unemployment_Rate, `County_Non-Hispanic_White_pct`,County_Hispanic_Native_American_pct, 
         County_Hispanic_White_pct,`County_Non-Hispanic_Native_American_pct`, other_demo_groups_pct)
  
res_store_tbl <- round(cor(res_store_mnd),2)
res_store_tbl
```

**Comment**\
We see from the data following:\
From the regression table, we see that there are generally weak
correlations between total sales and the other variables in the data
set. It is only in relation to County_Unemployment_Rate that we can see
a connection, but it is quite weak, only 0.41.\
<br/>

### **Summary**

It is chosen to use week number instead of month number because it
corresponds to time data in other data sets.

It is also chosen to use per \$ 1000 for the variables Sale and Profit.

It has been chosen to group the data set into price groups to see where
the focus of Sales is.

We have shown that the total Sales are in the price group
`r sales_most_pricegroup_mnd[1,1]` with sales
\$`r sales_most_pricegroup_mnd[1,3]/1000`.
`r sales_most_pricegroup_mnd[1,4]` percent of sales is made up of this
price group.

The highest profit is in the price group
`r profit_share_most_pricegroup_mnd[1,1]` with
\$`r profit_share_most_pricegroup_mnd[1,5]/1000` . The share of the
total profit is `r profit_share_most_pricegroup_mnd[1,6]` percent. In
this price group, it is `r bestselling_pricegroup_item[1,1]` that has
the highest sales.

We see that for total Sales, it is `r sales_most_item_mnd[1,1]` that has
the highest sales with \$`r sales_most_item_mnd[1,4]/1000`.
`r sales_most_item_mnd[1,1]` has `r sales_most_item_mnd[1,5]` percent of
total sales. `r profit_most_item_mnd[1,1]` has
`r profit_most_item_mnd[1,7]` percent of the total profit.

We have looked at how the total sales are distributed among the stores
and find that this month, it is No `r Store_most_sales[1,1]`
`r Store_most_sales[1,2]` that has the highest sales with \$
`r Store_most_sales[1,5]`. This represents `r Store_most_sales[1,6]`
percent of total sales.

An assessment has been made of non-economic variables such as location
and drive-in solutions' possible impact on sales. It turns out that in
this dataset with this choice of time period, it is likely to matter.

A simple regression has also been carried out which shows that there is
generally no connection between demographic variables and sales. The
same applies to crime rates. There is a slight positive correlation
between sales and the unemployment rate. This only applies to data in
this data set as choosing a different time period may give a different
result.

```{=html}
<pre>

</pre>
```
## **Oppgave 4: Kan dataene brukes til å planlegge nye utsalg.**

The data base that we have reviewed may indicate that it is better to
establish stores outside a shopping center than inside. Together with
the drive-in solution, it seems to provide good conditions for success.
Besides this, there are so few connections between sales and other
variables that it is difficult to recommend that emphasis be placed on
an establishment strategy.
