---
title: "Semesteroppgave SOK-1005-v22"
author: "Oddmund Klæboe"
date: "18 5 2022"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

Loading packages:

```{r downloading packages}
library(tidyverse)
library(lubridate)
library(knitr)

```

### **Oppgave 1: Slå sammen alle datasettene**  




Tenkte å lage det slik at man kunne velge butikk/butikknr og velge
uke(r) for å få de tallene man ønsker  


**Downloading data**

```{r downloads}
file_location <- paste0(getwd(), '/')

AppWichStoreAttributes <-
  "https://data.mendeley.com/public-files/datasets/6htjnfs78b/files/26afd5e7-90b1-4301-ac5e-5905b38c4ec2/file_downloaded"
dest <- paste0(file_location, "AppWichStoreAttributes.csv")
download.file(AppWichStoreAttributes, dest)

county_crime <-
  "https://data.mendeley.com/public-files/datasets/6htjnfs78b/files/3691994e-2874-4ee6-b215-12e130c96175/file_downloaded"
dest <- paste0(file_location, "county_crime.csv")
download.file(county_crime, dest)

county_demographic <-
  "https://data.mendeley.com/public-files/datasets/6htjnfs78b/files/527e7486-7233-460a-99e0-3529b7cd7d49/file_downloaded"
dest <- paste0(file_location, "county_demographic.csv")
download.file(county_demographic, dest)

county_employment <-
  "https://data.mendeley.com/public-files/datasets/6htjnfs78b/files/846ac757-721e-4fd9-a414-5871f60de093/file_downloaded"
dest <- paste0(file_location, "county_employment.csv")
download.file(county_employment, dest)

WEEKLY_SALES_10STORES <-
  "https://data.mendeley.com/public-files/datasets/6htjnfs78b/files/b963fdd1-0df9-4257-bd62-65256ec9d57c/file_downloaded"
dest <- paste0(file_location, "WEEKLY_SALES_10STORES.csv")
download.file(WEEKLY_SALES_10STORES, dest)

WEEKLY_WEATHER <-
  "https://data.mendeley.com/public-files/datasets/6htjnfs78b/files/b8708b3c-c5e8-456e-83f4-9f23f472a840/file_downloaded"
dest <- paste0(file_location, "WEEKLY_WEATHER.csv")
download.file(WEEKLY_WEATHER, dest)

# Removes all noise since we have finished downloading
rm(list=ls()) 
```

**I change the names of the datatables:**\
- AppWichStoreAttributes = df1_attr\
- county_crime = df2_crime\
- county_demographic = df3_demo\
- county_employment = df4_em\
- weekly_weather = df5_weather\
- weekly_sales_10stores = df6_sales

**I go through the dfs and make necessary changes in names and formats:**


```{r renaming and going through the dfs}
# some changes
# change the names of Store_County and Store_Weather_Station
df1_attr <- AppWichStoreAttributes <- read_csv("AppWichStoreAttributes.csv")
df1_attr_2 <-df1_attr %>%
  rename(County_Name = Store_County, Weather_Station = Store_Weather_Station)
head(df1_attr_2)

# no changes:
df2_crime <- county_crime <- read_csv("county_crime.csv")
head(df2_crime)

# no changes:
df3_demo <- county_demographic <- read_csv("county_demographic.csv")
head(df3_demo)

# no changes:
df4_em <- county_employment <- read_csv("county_employment.csv")
head(df4_em)

# some changes:
# Formatting the weather date:
df5_weather <- WEEKLY_WEATHER <- read_csv("WEEKLY_WEATHER.csv")

df5_weather$Weather_Date <- 
  as.Date(df5_weather$Weather_Date, format = "%d/%m/%Y") 
head(df5_weather)

# Some changes:
# rename the store number column
# some dates are missing in the Date column, I use the other date columns.
# removing unnecessary columns
# changing the letters in Description (names of items)


df6_sales <- WEEKLY_SALES_10STORES <- read_csv("WEEKLY_SALES_10STORES.csv")
df6_sales_2 <- df6_sales %>% 
  rename(Store_Num = Store_num) %>% 
  mutate(Date = as.Date(with(df6_sales, paste(Year, Month, Day,sep="-")), "%Y-%m-%d"))

df6_sales_3 <- 
  subset (df6_sales_2, select = -c(Year, Month, Day)) 
df6_sales_3$Description = str_to_title(df6_sales_3$Description)
df6_sales_3

```


**I make one df out of alle df by using the function left_join():**

```{r merging dfs}
df1_df2 <- left_join(df1_attr_2, df5_weather, by = "Weather_Station")
df1_df3 <- left_join(df1_df2, df2_crime, by = "County_Name")
df1_df4 <- left_join(df1_df3, df3_demo, by = "County_Name")
df1_df5 <- left_join(df1_df4, df4_em, by = "County_Name")
All_df <- left_join(df1_df5, df6_sales_3, by = c("Store_Num", "Weather_Date" = "Date"))

# renaming weather_Date and filtering out Cost <= 0
All_df <- All_df %>% 
  rename("Date" = "Weather_Date") %>% 
  filter(Cost > 0) 
All_df

```

### **Oppgave 2: Lag en ukesrapport for et utsalg**

```{r links_to_name_and_period}
# Store Name
Store_No_Nm <- All_df %>% 
  filter(Store_Num == 14)  

# Week number
Week <- All_df %>% 
  filter(Weather_Week == 20) 
```

This is a weekly report for store no `r Store_No_Nm[1,2]` `r Store_No_Nm[1,1]` for the period week no `r Week[1,22]` which gives a overview and assessment of the results.  

A weekly sales report will be a status report of how the operations in a sale are going. Here, key issues will be how sales of various goods have been and on which goods the profit has been greatest.   
The dataset includes many types of items. It has therefore been chosen to divide the product range into price groups and look at which price group has the largest Sale and Profit. Here it is also looked at which item is sold the most within the price group with the largest Sales. It has also been chosen to look at goods with the lowest turnover, and this can be helpful in assessing whether to maintain the large breadth of the product range.   
It must be added that the Profit is Gross Profit so that there will not be a supplementary analysis.

```{r selecting variables for week report}
# selecting week and Store
sales_14_20 <-All_df %>% 
  filter(Weather_Week == 20, Store_Num == 14) %>% 
  select(Store_Name, Store_Num, Store_City, County_Name, Date, INV_NUMBER,Description, Price, Sold, 
         Sales, Tot_Sls, Unit_Cost, Cost,Cost_Percent, Margin, Profit) %>% 
  group_by(INV_NUMBER, Description, Price, Sold, Cost, Profit, Margin) %>% 
  summarise(Sales) %>% 
  ungroup()
```

```{r item most sales}
# Item that have most Sale
sales_most_item <-
  subset(sales_14_20, Sales = max(sales_14_20$Sales)) %>% 
  arrange(desc(Sales))
sales_most_item

```

```{r item most profitable}
Profit_most_item <-
subset(sales_14_20, Profit = max(sales_14_20$Profit)) %>% 
  arrange(desc(Profit))
Profit_most_item
```

```{r comparing two weeks}
# selecting the week before to compare development in Sales and Profit:
sales_14_19 <-All_df %>% 
  filter(Weather_Week == 19, Store_Num == 14) %>% 
  select(Store_Name, Store_Num, Store_City, County_Name, Date, INV_NUMBER,Description, Price, Sold, 
         Sales, Tot_Sls, Unit_Cost, Cost,Cost_Percent, Margin, Profit) %>%
  group_by(INV_NUMBER, Description, Price, Sold, Cost, Profit, Margin) %>% 
  summarise(Sales) %>% 
  ungroup()

# sum of variables:
sum(sales_14_19$Sales)
sum(sales_14_20$Sales)
sales_change = round(100 * ((sum(sales_14_20$Sales)/sum(sales_14_19$Sales)-1)),1)
sales_change

sum(sales_14_19$Profit)
sum(sales_14_20$Profit)
profit_change = round(100 * ((sum(sales_14_20$Profit)/sum(sales_14_19$Profit)-1)),1)
profit_change
```

**Comment**  
  
Total revenues of Sales for store `r Store_No_Nm[1,2]` `r Store_No_Nm[1,1]`  i week `r Week[1,22]` were \$ `r sum(sales_14_20$Sales)`. The total sales changed `r sales_change` percent compared to the week before.  
Total profit in the same period were \$ `r sum(sales_14_20$Profit)`. The change in total profit from the week before were  `r profit_change` percent.   
Sales revenues were greatest for `r sales_most_item[1,2]`. Total sales for this were \$ `r sales_most_item[1,8]`.
The most profitable item were `r Profit_most_item[1,2]`. For this, the total profit were \$ `r Profit_most_item[1,6]`.  

  
  
  


#### **Sorting the goods after price groups:**

Due to the large breadth of the product range, it has been chosen to take the product prices as a starting point for the analysis and group them. They are sorted in following groups. Price below \$1(here are also refund etc that is negative figures), between \$1 and \$2, and so on. The last price group is price over \$8.

```{r sorting price in groups}
sales_price_gr <- sales_14_20 %>% 
  group_by(price_group = ifelse(Price <= 1.0, "price_below_$1", 
                         ifelse(Price > 1 & Price <= 2, "price_$1", 
                         ifelse(Price > 2 & Price <= 3, "price_$2", 
                         ifelse(Price > 3 & Price <= 4, "price_$3", 
                         ifelse(Price > 4 & Price <= 5, "price_$4", 
                         ifelse(Price > 5 & Price <= 6, "price_$5", 
                         ifelse(Price > 6 & Price <= 7, "price_$6", 
                         ifelse(Price > 7 & Price <= 8.0, "price_$7", "price_over_$8"))))))))) %>% 
  summarise(Sold, Price, Sales, Profit)

# we remove all negative values from the dataset
sales_price_gr <-
  sales_price_gr %>%
  select(price_group, Sold, Price, Sales, Profit) %>%
  filter(price_group >= 0) %>%
  group_by(price_group) %>%
  summarise(Tot_sales = sum(Sales), Tot_sold = sum(Sold), Tot_profit = sum(Profit)) 

# Total sales and profit:
sum(sales_price_gr$Tot_sales)
max(sales_price_gr$Tot_sales)  

sum(sales_price_gr$Tot_profit)
max(sales_price_gr$Tot_profit)


sales_most <- 
  subset(sales_price_gr, Tot_sales == max(sales_price_gr$Tot_sales))


profit_most <- 
  subset(sales_price_gr, Tot_profit == max(sales_price_gr$Tot_profit))
```


#### **Share of sales and profit per pricegroup:**

```{r shares of sales and profit}
# share of sales and profit per pricegroup:
sales_share <-
  sales_price_gr %>% 
  mutate(Share_sales = round(100 *(sales_price_gr$Tot_sales/sum(sales_price_gr$Tot_sales)), 1),
         Share_profit = round(100 *(sales_price_gr$Tot_profit/sum(sales_price_gr$Tot_profit)),1)) %>% 
  select(price_group, Tot_sold, Tot_sales, Share_sales, Tot_profit, Share_profit)

Shares_most <- 
  subset(sales_share, Share_sales == max(sales_share$Share_sales)) 
```

```{r table 1 Amount and shares for Sale and Profit}
Goods_shares <- sales_share[1:9, ]
  knitr::kable(Goods_shares, 
               caption = "Table 1: Amount and shares for Sale and Profit")
```
**Comment**  
We see from table 1 the highest revenue comes from sale in the pricegroup `r Shares_most[1,1]`. The sales here are \$ `r Shares_most[1,3]` and profit are \$ `r Shares_most[1,5]`. This pricegoup has a share of `r Shares_most[1,4]` percent of total Sale and `r Shares_most[1,6]` percent of total Profit.



#### **Making a plot of sales per pricegroup:**

```{r figure 1 sales salsgroup}
figure_1 <-
sales_price_gr %>% 
ggplot(aes(x=price_group, y = Tot_sales))+
  geom_bar(stat= "identity", fill = "steelblue") +
  geom_text(aes(label=Tot_sales), vjust= -0.3, size=3.5)+
  labs(title = "Figure 1: Sales per pricegroup", x = "Pricegroups in $", y = "Total Sales per pricegoup",
        caption = "Figure 1 shows how Sales are distributed across pricegroups. For example' price_$5 mean goods that cost from $4 to $5.") +
  theme_classic()
figure_1

```

**Comment**  

We see from Figure 1 that the highest sales are of goods in the price group `r sales_most[1,1]`. Total Sales are \$`r sales_most[1,2]`. The first price group has negative sales figures. The reason for this is that it also contains refunds etc.  


#### **Making a plot of shares of sales per pricegroup:**

```{r figure 1a share of sales}
figure_1a <-
  sales_share %>% 
  ggplot(aes(x=price_group, y = Share_sales))+
  geom_bar(stat= "identity", fill = "steelblue") +
  geom_text(aes(label=Share_sales), vjust= -0.3, size=3.5)+
  labs(title = "Figure 1a: Shares of Sales per pricegroup", x = "Pricegroups in $", y = "Share of Sales per pricegoup in percent",
       caption = "Figure 1 shows how shares of Sales are distributed across pricegroups.") +
  theme_classic()
figure_1a
```

**Comment**  
We see from Figure 1a that the highest share of sales are of goods in the price group `r Shares_most[1,1]`. This price group has `r Shares_most[1,4]` per cent of all sales.  


#### **Making a plot of profit per pricegroup:**

```{r figure 2 profit}
figure_2 <-
  sales_price_gr %>% 
  ggplot(aes(x=price_group, y = Tot_profit))+
  geom_bar(stat= "identity", fill = "steelblue") +
  geom_text(aes(label=Tot_profit), vjust= -0.3, size=3.5)+
  labs(title = "Figure 2: Profit per pricegroup", x = "Pricegroups in $", y = "Total Profit per pricegoup",
       caption = "Figure 2 shows how Profit are distributed across pricegroups. For example price_$5 mean goods that cost from $4 to $5.") +
  theme_classic()
figure_2
```

**Comment**  
We see from Figure 2 that the largest profit is achieved from sales in the price group `r profit_most [1,1]`. The total profit here is \$ `r profit_most [1,4]`. The first price group is negative. The reason for this is that it also contains refunds etc. who will negatively affect profits.  
  
  





#### **The best-selling items**


```{r best selling}
# Highest number of items sold
Goods_most<- sales_14_20 %>% 
  filter(Sold > 10) %>% 
  arrange(desc(Sold))

# Higest revenue from Sales
Goods_most_sales<- sales_14_20 %>% 
  filter(Sold > 10) %>% 
  arrange(desc(Sales))

# Higest Profit
Goods_most_p <- sales_14_20 %>% 
  filter(Sold > 10) %>% 
  arrange(desc(Profit))

sum(Goods_most$Sold)
sum(Goods_most$Sales)
sum(Goods_most$Profit)

```

We make a table of most sold items

```{r most sold table 2}
Goods_most <- Goods_most[1:10, ]
  kable(Goods_most, caption = "Table 2: Most sold items")
```

**Comment**  
Table 2 shows the 10 items that sell the most in Store `r Store_No_Nm[1,2]` `r Store_No_Nm[1,1]` in this week. Sales for these 10 are total $ `r sum(Goods_most$Sales)` and account for `r round(100 *(sum(Goods_most$Sales)/sum(sales_14_20$Sales)),1)` percent of total sales.

It appears from Table 2 that the mostselling item were `r Goods_most [1,2]`.  `r Goods_most [1,4]` units of `r Goods_most [1,2]` were sold. That account for $`r Goods_most[1,8]` in sale. The sale of `r Goods_most[1,2]` accounts for `r round(100 *(Goods_most[1,8]/sum(sales_14_20$Sales)),1)` percent of total Sales.

In second place we find `r Goods_most[2,2]` with `r Goods_most[2,4]` units and $ `r Goods_most[2,8]` of sale. As third ranked, we find `r Goods_most[3,2]` with `r Goods_most[3,4]` units and $ `r Goods_most[3,8]` of sale.

**Most profitable items**  

`r Goods_most_p[1,2]` is the most profitable sold item. The profit is \$ `r Goods_most[1,6]`. This amounts to
`r round(100 *(Goods_most_p[1,6]/sum(sales_14_20$Profit)),1)` percent of total profit.

**The lowest sales**

```{r low sales}
# lowest number of items sold
Goods_low<- sales_14_20 %>% 
  filter(Sold > 10, Price > 0) %>% 
  arrange(Sold)

# lowest revenue from Sales
Goods_low_sales<- sales_14_20 %>% 
  filter(Sold > 10, Price > 0) %>% 
  arrange(Sales)

# lowest Profit
Goods_lowest_p <- sales_14_20 %>% 
  filter(Sold > 10, Price > 0) %>% 
  arrange(Profit)
Goods_lowest_p
```

**We make a table of the lowest number of sold items**

```{r low sales table 3}
Goods_low <- Goods_low[1:10, ]
  kable(Goods_low, caption = "Table 3: lowest number sold items")
```
**Comment**  
From table 3 we see that there are many items that not are selling good.  
  
  

##### **Summary:**  
The week sales report for Store `r Store_No_Nm[1,2]` `r Store_No_Nm[1,1]` in week `r Week[1,22]`show that total revenues for 
were \$ `r sum(sales_14_20$Sales)`, and total Profit were \$ `r sum(sales_14_20$Profit)`.  
The most-selling item were `r sales_most_item[1,2]`. Total sales for this were \$ `r sales_most_item[1,8]`. The total sales changed `r sales_change` percent compared to the week before.  
The most profitable item were `r Profit_most_item[1,2]`.
For this, the total profit were \$ `r Profit_most_item[1,6]`. The change in total profit from the week before were \ `r profit_change` percent.  




### **Oppgave 3: Lag en månedlig salgrapport på aggregert nivå.**

Månedsdata - lage data for alle mnd slik at man kan se utvikling, totalt
og hvert utsalg. Grafikk

Omsetning totalt utvikling(Sales) Grafikk

Deler inn i prisgrupper

Hvilke prisgrupper gir mest/minst profitt

Omsetning pr prisgruppe Grafikk

Andel av omsetning

Hvor mye utgjør hver prisgruppe av salget. Grafikk

Hvem er kunder og hva spiser de (koble demogrupper og varegrupper) - slå
sammen

Belyse om det er sammenheng mellom ledighet i området og omsetning
Belyse om det er sammenheng mellom demo i området og omsetning Belyse om
det er sammenheng mellom crime i området og omsetning

#### **Introduction**  
In a monthly report, the group management will be able to see the total results of the business and make strategic assessments based on this. This means that Sales and Profits in total and per Store are central. On the basis of this, it will be possible to assess whether special measures are necessary for some Stores. Comparison of data on the same item across Stores will also be relevant. This can form the basis for looking at whether the product range should be the same in all Stores.  

It has been chosen here to divide the product range into price groups and look at which price group has the largest Sales and Profits. Here it is also looked at which item is sold the most within the price group with the largest Sales.  

It has also been chosen to look at items with the lowest sales, which can be helpful in assessing whether to maintain the large breadth of the product range.  

The dataset is not very suitable for creating real monthly data because it is based on weeks. It will be the case that a week can contain data for several months. It is therefore recommended that the group switch to a day-based accounting system so that monthly data can be created in a simpler way.



**Cleaning the data set for unnecessary variable and grouping some
variables:**

```{r cleaning of df}

All_df_rev <- All_df %>% 
  
# We merge small population groups into one variable:
  mutate("All_Other_Groups"= rowSums(All_df[ , c("County_Non-Hispanic_Black","County_Non-Hispanic_Asian",
                                                 "County_Non-Hispanic_Pacific_Islander", "County_Non-Hispanic_Two_or_more",
                                                 "County_Hispanic_Black", "County_Hispanic_Asian",
                                                 "County_Hispanic_Pacific_Islander", "County_Hispanic_Two_or_more")])) %>%  
  
# and select the relevant variables:    
  select(Store_Name, Store_Num, Store_City, County_Name, Weather_Station, Date, Weather_Week,    
         Description, Price, Sold,Sales, Tot_Sls, Cost, Margin, Profit, Annual_Rent_Estimate,Store_Location, Store_Drive_Through, 
         Store_Competition_Fastfood, Weather_Bad_Weather_days, County_Total_Crime_Rate, County_Violent_Rate, 
         County_Property_Rate, County_Society_Rate, County_Unemployment_Rate, `County_Non-Hispanic_White`,
         `County_Non-Hispanic_Native_American`, County_Hispanic_White,  County_Hispanic_Native_American, All_Other_Groups, 
         County_Total_Census_Pop, ) %>% 
  
  
# making percentages of population groups
  mutate(All_Other_Groups_pct = round(100 * (All_Other_Groups/County_Total_Census_Pop),1),
         `County_Non-Hispanic_White_pct` = round(100 * (`County_Non-Hispanic_White`/County_Total_Census_Pop),1),
         `County_Non-Hispanic_Native_American_pct` = round(100 * (`County_Non-Hispanic_Native_American`/`County_Total_Census_Pop`),1),
         `County_Hispanic_White_pct` = round(100 * (`County_Hispanic_White`/County_Total_Census_Pop),1),
         `County_Hispanic_Native_American_pct` = round(100 * (`County_Hispanic_Native_American`/County_Total_Census_Pop),1)) %>% 
  
# removing the variables we dont use
  select(- c(`County_Non-Hispanic_White`,
             `County_Non-Hispanic_Native_American`, County_Hispanic_White,  County_Hispanic_Native_American, All_Other_Groups, 
             County_Total_Census_Pop)) %>% 
  rename(Week_No = Weather_Week) 


All_df_rev$County_Violent_Rate <-
  round(All_df_rev$County_Violent_Rate,1)
```

**We choose 4 weeks from 40 to 43:**

```{r choosing weeks}
# choosing weeks
sales_mnd <-All_df_rev %>% 
  filter(Week_No >= 40, Week_No <= 43) %>% 
  select(Store_Name, Store_Num, Store_City, County_Name, Date,Description, Price, Sold, 
         Sales, Profit) %>% 
  group_by(Description, Price, Sold, Sales,Profit) %>% 
  summarise(Sold, Sales) %>% 
  ungroup()
sales_mnd

# sum of variables:
sum(sales_mnd$Sales)
sum(sales_mnd$Profit)
```
**Comment**  
The total Sales revenue from all the Stores in this month were $`r (sum(sales_mnd$Sales))/1000`. Total Profit amounts to $`r (sum(sales_mnd$Profit))/1000`. All numbers are in thousands of $.


**Sorting the goods after price groups:**  
Due to the large breadth of the product range, it has been chosen to take the product prices as a starting point for the analysis and group them. They are sorted in following groups. Price below \$1(here are also refund etc that is negative figures), between \$1 and \$2, and so on. The last price group is price over \$8.

```{r sorting pricegroups for use on aggregate level}
sales_price_gr_mnd <- sales_mnd %>% 
  group_by(price_group = ifelse(Price <= 1.0, "price_$1", 
                         ifelse(Price > 1 & Price <= 2, "price_$2", 
                         ifelse(Price > 2 & Price <= 3, "price_$3", 
                         ifelse(Price > 3 & Price <= 4, "price_$4", 
                         ifelse(Price > 4 & Price <= 5, "price_$5", 
                         ifelse(Price > 5 & Price <= 6, "price_$6", 
                         ifelse(Price > 6 & Price <= 7, "price_$7", 
                         ifelse(Price > 7 & Price <= 8.0, "price_$8", "price_over_$8"))))))))) %>% 
  summarise(Description, Sold, Price, Sales, Profit)
sales_price_gr_mnd
```

grouping after pricegroups:

```{r grouping by pricegroups}
sales_price_gr_mnd <-
  sales_price_gr_mnd %>%
  select(Description, price_group, Sold, Price, Sales, Profit) %>%
  filter(price_group >= 0) %>%
  group_by(Description, price_group) %>%
  summarise(Tot_sales = sum(Sales), Tot_sold = sum(Sold), Tot_profit = sum(Profit)) 
sales_price_gr_mnd
```

```{r}
# Total sales and profit:
sum(sales_price_gr_mnd$Tot_sales)


max(sales_price_gr_mnd$Tot_sales) 

sum(sales_price_gr_mnd$Tot_profit)
max(sales_price_gr_mnd$Tot_profit)
```



**Amount and Share of sales and profit per pricegroup:**

```{r Amount of Sales and Profit}
# Amount of Sales
sales_most_mnd <- 
  subset(sales_price_gr_mnd, Tot_sales == max(sales_price_gr_mnd$Tot_sales))
sales_most_mnd

# Amount of Profit
profit_most_mnd <- 
  subset(sales_price_gr_mnd, Tot_profit == max(sales_price_gr_mnd$Tot_profit))
profit_most_mnd

```
**Comment**  
The highest revenue from Sales this month was from `r sales_most_mnd[1,1]`. This amounted to $`r (sales_most_mnd[1,3])/1000` in revenue. The largest profit was from `r profit_most_mnd[1,1]`. Here, the sum was $ `r (profit_most_mnd[1,5])/1000`. All numbers are in thousands of $.


```{r amount and shares of Sales and Profit}
amount_share_mnd <-
  sales_price_gr_mnd %>% 
  mutate(Share_sales_mnd = round(100 *(Tot_sales/sum(sales_price_gr_mnd$Tot_sales)), 1),
         Share_profit_mnd = round(100 *(Tot_profit/sum(sales_price_gr_mnd$Tot_profit)),1)) %>% 
  select(Description, price_group, Tot_sold, Tot_sales, Share_sales_mnd, Tot_profit, Share_profit_mnd) %>% 
  arrange(desc(Tot_sales))
amount_share_mnd

sales_most_mnd <- 
  subset(amount_share_mnd, Tot_sales == max(amount_share_mnd$Tot_sales))
sales_most_mnd

Shares_most_mnd <- 
  subset(amount_share_mnd, Share_sales_mnd == max(amount_share_mnd$Share_sales_mnd)) 
Shares_most_mnd  
```

We make a table of amount and shares

```{r table 4 amount and shares}
amount_shares_mnd <- amount_share_mnd[1:10, ]
 kable(amount_shares_mnd, caption = "Table 4: Amount and shares for Sale and Profit")

```

**Comment**  
We see from Table 4 that the highest sales revenue comes from the `r amount_shares_mnd [1,1]`.   

In this price group, goods were sold for a total of \$ `r (sales_most_mnd [1,4])/1000`. The profit from sales of this item  was `r (sales_most_mnd [1,6])/1000`. All numbers are in thousands of $.



**Making a plots of sales of mostselling items:**

```{r plot figure 3 sales revenue per mostselling items}
# Sales pr item
figure_3 <-
  amount_shares_mnd %>% 
  ggplot(aes(x= Description, y = Tot_sales))+
  geom_bar(stat= "identity", fill = "steelblue") +
  geom_text(aes(label=Tot_sales/1000), vjust= -0.3, size=3.5)+
  labs(title = "Figure 3: Sales revenue of mostselling items in one month", 
       x = "Item Description", y = "Sales revenue in $1000",
       caption = "Figure 3 shows total revenues from the 10 mostselling items.") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
figure_3
```

**Comment**  
We see from Figure 3 that the largest Sales revenue is achieved by `r sales_most_mnd [1,1]`. The sales revenue from this item amounts to $`r (Shares_most_mnd[1,4])/1000`. Numbers are in thousands of $.   

```{r plot figure 3 mostselling items}
# Sales as shares of total
figure_3a <-
  amount_shares_mnd %>% 
  ggplot(aes(x= Description, y = Share_sales_mnd))+
  geom_bar(stat= "identity", fill = "steelblue") +
  geom_text(aes(label=Share_sales_mnd), vjust= -0.3, size=3.5)+
  labs(title = "Figure 3a: Sales of mostselling items as shares of total Sales in one month", 
       x = "Item Description", y = "Sales shares per item",
       caption = "Figure 3 shows the 10 mostselling item as shares of total Sale .") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
figure_3a
```

**Comment**   
We see from Figure 3a that the largest share of total Sales is achieved by `r sales_most_mnd [1,1]`. This item has `r Shares_most_mnd [1,5]` percent of total Sales.


  
**Making a plot of shares of sales per pricegroup:**



**Comment**  
We see from figure 3a that the largest share of sales in this month are in the price group `r Shares_most_mnd[1,1]`. This price group have `r Shares_most_mnd [1,4]` percent of all Sales.  



**Making a plot of profit per pricegroup:**

```{r plot profit per pricegroup}
figure_4 <-
  amount_shares_mnd %>% 
  ggplot(aes(x= Description, y = Share_profit_mnd))+
  geom_bar(stat= "identity", fill = "steelblue") +
  geom_text(aes(label= Share_profit_mnd), vjust= -0.3, size=3.5)+
  labs(title = "Figure 4: Profits of mostselling items as shares of total Profit in one month", 
       x = "Item Description", y = "Profit shares per item",
       caption = "Figure 3 shows the 10 mostselling item as shares of total Profit .") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
figure_4
```

**Comment**

We see from Figure 4 that the largest profit is achieved from sales in the price group `r profit_most_mnd[1,1]`. The total profit here is \$ `r (profit_most_mnd[1,4])/1000`. The first price group is negative. The reason for this is that it also contains refunds etc. who will negatively affect profits. All numbers are in thousands of $.

#### **The best-selling items in**

Hvilke varer finner vi i prisgruppen med størst omsetning. Vi finner
(lenke til varer) i denne gruppen.

```{r most selling good mnd}
Goods_most_mnd<- sales_mnd %>%
  filter(Price > 4 & Price <= 5, Sold > 10) %>% 
  arrange(desc(Sales))

sum(Goods_most_mnd$Sold)
sum(Goods_most_mnd$Sales)
sum(Goods_most_mnd$Profit)
```

We make a table of most sold goods NB må ordnes med sumtall for hvert
vareslag

```{r table 4 most sold goods mnd}
Goods_most_mnd <- Goods_most_mnd[1:10, ]
  kable(Goods_most_mnd, caption = "Table 4: Most sold goods in month")
```
